Using Cell Phone Keyboards is (\NP) Hard

### 2010-01-20 article.tex
Trying for a new aticle.tex with \Question and \Instance defined

\newcommand{\newword}[1]{{\bf #1}}
\newcommand{\NP}{\ensuremath{\mathcal{NP}}}
\spnewtheorem{prob}[theorem]{Problem}{\bfseries}{\itshape}

@@

Before we get any farther, let us sketch the basic problem that we will keep
revising and revisiting throughout this paper.
\begin{prob}[{\sc MinimumKeystrokes}]~\\
{\sc Instance}: A set of letters corresponding to an alphabet $A$ $(|A| =
n)$, a number of keys $k$, an input method $\mathrm{IN}$, and a set of
tuples of words and frequencies $W$.  The frequencies in $W$ are integers,
and the words are made up of solely of elements of $A$.  We will
treat $IN$ as a function which, given a partition of $A$ and a word
$w$, returns how many keystrokes are required to type $w$.~\\
        ~\\
{\sc Question}: What is the best partition of $A$ into $k$ sets, such that the
total number of keystrokes to type every word in $W$ its associated frequency
times is minimized?  Equivalently, what is the partition of $P$ of $A$ ($|P| = k$) that
will minimize
$$\sum_{(w,f)\in W} f*\mathrm{IN}(w,P)$$
\label{probtemplate}
\end{prob}

@@

\begin{prob}[{\sc BasicCellPhoneTyping}]~\\
{\sc Instance}: A set of letters corresponding to an alphabet $A$ $(|A| =
n)$, a number of keys $k$, and a set of
tuples of words and frequencies $W$.  The frequencies in $W$ are integers,
and the words are made up of solely of elements of $A$. \\
        ~\\
{\sc Question}: What is the best partition $P$ of $A$ into $k$ sequences, such
that the total number of button presses to type every word in $W$ its
associated frequency times is minimized?  The number of button presses is equal
to the order of the letter in the sequence assigned to a given key.  For
example, in the key $2\to\{abc\}$, $a$ requires one keystroke and $c$ requires
3.  Equivalently, if we denote the number of button presses required to type
letter $a$ with partition $P$ as $\mathrm{IN}_P(a)$, then we want to find the $P$ that
minimizes $$\sum_{(w,f) \in W}\sum_{c\in w} \mathrm{IN}_P(c) * f$$
\label{bcpt}
\end{prob}

@@

\begin{prob}[{\sc MinimumT9Keystrokes}]~\\
\label{thm:minstrokes}
{\sc Instance} An alphabet $A$, a set of words and their associated frequencies $W$, and a number of keys $k$ ($|A| > k$).\\
~\\
{\sc Question} What is the partion $P$ of $A$ into $k$ sets which minimizes the total button presses required to enter the entire corpus using the T9 input method?
\end{prob}

@@

\begin{prob}~\\
{\sc Instance} An alphabet $A$, a set of words and their associated frequencies $W$, a number of keys $k$ ($|A| > k$), and a number $t$.\\
~\\
{\sc Question} Is there a partition of $A$ into $k$ sets in which the total number of keystrokes required to enter the entire corpus using the T9 input method requires no more than $t$ button presses?
\end{prob}

Before proceeding, we note that this we note that this problem is in \NP, as
any assignment may be verified to have and expected number of characters of
less than $f$ in time proportional to the total length of all the words in $W$.

It is highly likely this problem is \NP-complete, due to a reduction from one of the optimal trie problems.  But I'm not there yet.

However, our arbitrary remappings of the keyboard may be quickly deemed unusable.  Therefore, we define a refinement of the problem, in which the characters are required to be kept in alphabetical order as a sequence, and our only choices are about how to divide the subsequence into key assignments.  The default layout can be described as the partition $abc|def|ghi|jkl|mno|pqrs|tuv|wxyz$.  Is this partition optimal?  In order to decide this, we define the problem {\sc MinimumKeystrokePartition} as:
\begin{prob}[{\sc MinimumKeystrokePartition}]~\\
\label{thm:minpartition}
{\sc Instance} A sequence of letters $A = \{a_1, a_2, \ldots \}$, a set of words and their associated probabilities $W$, and a set of keys $K = \{2, 3, 4, \ldots \}$ ($|A| > |K|$).\\
~\\
{\sc Question} What is the mapping $f$ of $A \to K$ which minimizes the expected number of characters to type a word from $W$, and where $f(a_1) = 2$ and if $f(a_i) = k$, then either $f(a_{i+1}) = k$ or $f(a_{i+1}) = k+1$?
\end{prob}

As before, the corresponding decision problem is
\begin{prob}~\\
{\sc Instance} A sequence of letters $A = \{a_1, a_2, \ldots \}$, a set of words and their associated probabilities $W$, a set of keys $K = \{2, 3, 4, \ldots \}$ ($|A| > |K|$), and a parameter $e$.\\
~\\
{\sc Question} Is there a mapping $f$ of $A \to K$ in which the expected number of characters to type a word from $W$ is less than $e$, and where $f(a_1) = 2$ and if $f(a_i) = k$, then $f(a_{i+1}) \in \{k, k+1\}$?
\end{prob}

>>>>

\newcommand{\newword}[1]{{\bf #1}}
\newcommand{\NP}{\ensuremath{\mathcal{NP}}}
\newcommand{\Instance}{\Instance}
\newcommand{\Question}{~\\
\Question}
\spnewtheorem{prob}[theorem]{Problem}{\bfseries}{\itshape}

@@

Before we get any farther, let us sketch the basic problem that we will keep
revising and revisiting throughout this paper.
\begin{prob}[{\sc MinimumKeystrokes}]~\\
\Instance\ A set of letters corresponding to an alphabet $A$ $(|A| =
n)$, a number of keys $k$, an input method $\mathrm{IN}$, and a set of
tuples of words and frequencies $W$.  The frequencies in $W$ are integers,
and the words are made up of solely of elements of $A$.  We will
treat $IN$ as a function which, given a partition of $A$ and a word
$w$, returns how many keystrokes are required to type $w$.

\Question\ What is the best partition of $A$ into $k$ sets, such that the
total number of keystrokes to type every word in $W$ its associated frequency
times is minimized?  Equivalently, what is the partition of $P$ of $A$ ($|P| = k$) that
will minimize
$$\sum_{(w,f)\in W} f*\mathrm{IN}(w,P)$$
\label{probtemplate}
\end{prob}

@@

\begin{prob}[{\sc BasicCellPhoneTyping}]~\\
\Instance\ A set of letters corresponding to an alphabet $A$ $(|A| =
n)$, a number of keys $k$, and a set of
tuples of words and frequencies $W$.  The frequencies in $W$ are integers,
and the words are made up of solely of elements of $A$. \\
        ~\\
\Question\ What is the best partition $P$ of $A$ into $k$ sequences, such
that the total number of button presses to type every word in $W$ its
associated frequency times is minimized?  The number of button presses is equal
to the order of the letter in the sequence assigned to a given key.  For
example, in the key $2\to\{abc\}$, $a$ requires one keystroke and $c$ requires
3.  Equivalently, if we denote the number of button presses required to type
letter $a$ with partition $P$ as $\mathrm{IN}_P(a)$, then we want to find the $P$ that
minimizes $$\sum_{(w,f) \in W}\sum_{c\in w} \mathrm{IN}_P(c) * f$$
\label{bcpt}
\end{prob}

@@

\begin{prob}[{\sc MinimumT9Keystrokes}]~\\
\label{thm:minstrokes}
\Instance\ An alphabet $A$, a set of words and their associated frequencies $W$, and a number of keys $k$ ($|A| > k$).\\
~\\
\Question\ What is the partion $P$ of $A$ into $k$ sets which minimizes the total button presses required to enter the entire corpus using the T9 input method?
\end{prob}

@@

\begin{prob}~\\
\Instance\ An alphabet $A$, a set of words and their associated frequencies $W$, a number of keys $k$ ($|A| > k$), and a number $t$.\\
~\\
\Question\ Is there a partition of $A$ into $k$ sets in which the total number of keystrokes required to enter the entire corpus using the T9 input method requires no more than $t$ button presses?
\end{prob}

Before proceeding, we note that this we note that this problem is in \NP, as
any assignment may be verified to have and expected number of characters of
less than $f$ in time proportional to the total length of all the words in $W$.
In order to prove completeness, however, we first prove the \NP-completeness of an intermediate problem: {\sc UniquePathColoring}.

\begin{prob}[{\sc UniquePathColoring}]~\\
\label{upcolor}
\Instance\ A graph $G=(V,E)$, a set of paths $P$, and a parameter $k$.\\
~\\
\Question\
\end{prob}

However, our arbitrary remappings of the keyboard may be quickly deemed unusable.  Therefore, we define a refinement of the problem, in which the characters are required to be kept in alphabetical order as a sequence, and our only choices are about how to divide the subsequence into key assignments.  The default layout can be described as the partition $abc|def|ghi|jkl|mno|pqrs|tuv|wxyz$.  Is this partition optimal?  In order to decide this, we define the problem {\sc MinimumKeystrokePartition} as:
\begin{prob}[{\sc MinimumKeystrokePartition}]~\\
\label{thm:minpartition}
\Instance\ A sequence of letters $A = \{a_1, a_2, \ldots \}$, a set of words and their associated probabilities $W$, and a set of keys $K = \{2, 3, 4, \ldots \}$ ($|A| > |K|$).

\Question\ What is the mapping $f$ of $A \to K$ which minimizes the expected number of characters to type a word from $W$, and where $f(a_1) = 2$ and if $f(a_i) = k$, then either $f(a_{i+1}) = k$ or $f(a_{i+1}) = k+1$?
\end{prob}

As before, the corresponding decision problem is
\begin{prob}~\\
\Instance\ A sequence of letters $A = \{a_1, a_2, \ldots \}$, a set of words and their associated probabilities $W$, a set of keys $K = \{2, 3, 4, \ldots \}$ ($|A| > |K|$), and a parameter $e$.

\Question\ Is there a mapping $f$ of $A \to K$ in which the expected number of characters to type a word from $W$ is less than $e$, and where $f(a_1) = 2$ and if $f(a_i) = k$, then $f(a_{i+1}) \in \{k, k+1\}$?
\end{prob}
